Light versus dark in strong-lens galaxies: Dark matter haloes that are rounder than their stars

### 2015-09-18 shapes/paper.tex
Small changes to the abstract by Justin

\begin{abstract}
We measure the shape and alignment of the stellar and dark matter mass distribution in 11 strong-lens galaxies. We show that the projected dark matter density profile in these systems -- under the assumption of a Chabrier stellar initial mass function -- shows significant variation from galaxy to galaxy. Those with an outermost image beyond $\sim 10$\,kpc are very well fit by a projected NFW profile; those with images within 10\,kpc appear to be more concentrated than NFW, as expected if their dark halos contract due to baryonic cooling. We find that over several half-light radii, the dark matter haloes of these lenses are rounder than their stellar mass distributions. While the haloes are never more elliptical than $s_{dm} = 0.3$ (where $s$ is defined such that $s=0$ is spherical and $s=1$ is infinitely flattened), their stars can extend to $s_* > 0.3$. Galaxies with high dark matter ellipticity and weak external shear show strong alignment between light and dark; those with strong shear ($\gamma \gtrsim 0.1$) can be highly misaligned. This is reassuring since isolated misaligned galaxies are expected to be unstable. Our results provide a new constraint on galaxy formation models. For a given cosmology, these must explain the origin of both very round dark matter haloes and misaligned strong-lens systems.
{\bf JR. I hope you agree with my analysis of the radial profiles. I've spotted a very cute little result here already and thought we should highlight it. I discuss it further in the results and conclusions. Also, we now define $s$ to be the `ellipticity'. But later on when discussing weak lensing we define a different ellipticity parameter $e$. The referee took us to task on this also I think. Are there really two different definitions of ellipticity in the literature? If so, perhaps we should clarify this up-front when we define $s$. Alternatively, if $s$ is unique to our paper then why not just change $s \rightarrow e$ throughout? Or we could call our $s$ parameter the `shape parameter' and $e$ the `ellipticity'. I have no strong view on which is best, but it's a bit confusing as it is right now}.
\end{abstract}

@@

Our lens sample is small -- just 11 strong lensing systems -- yet it is interesting that we find no lens with strong misalignment and weak external shear. This is perhaps to be expected. \citet{XXX} show that YYY.

\textbf{JR. I think comparisons with weak lensing have a section of their own and don't need to go in the conclusions. I moved them to this discussion section, though, to ensure that they stand out. I added a new subsection in discussion also about misalignments and stability (above). This should deal with that issue.}
%Relevant references for misalignments being unstable: 
%- http://adsabs.harvard.edu/abs/1978MNRAS.183..779B : 1978MNRAS.183..779B
%- http://adsabs.harvard.edu/abs/1981MNRAS.196..455B : 1981MNRAS.196..455B
%- http://adsabs.harvard.edu/abs/1979ApJ...233..872H : 1979ApJ...233..872H
%- http://adsabs.harvard.edu/abs/1984ApJ...276..101S : 1984ApJ...276..101S
%- http://adsabs.harvard.edu/abs/1987A%26A...182..202R : 1987A&A...182..202R
%- http://adsabs.harvard.edu/abs/1988A%26A...206..269M : 1988A&A...206..269M
%- http://adsabs.harvard.edu/abs/2007ApJ...670.1027A : 2007ApJ...670.1027A
% AND :: 1979ApJ...233..872H,1988A&A...206..269M,2007ApJ...670.1027A,2015arXiv150203429D

\section{Conclusions}\label{sec:conclusions}
We have measured the radial density profile, shape and alignment of the stellar and dark matter distributions in 11 lens systems using a new non-parametric lens tool, \Glass. We focussed on lenses that have either time delay data or stellar mass maps that contribute significantly to the central potential. In a previous paper, we showed that data of this quality are required to determine the projected shape of dark matter haloes \citep{2014MNRAS.445.2181C}. We measured the shape and alignment using the eigenvalues $\lambda_i$ and eigenvectors of the 2D moment of inertia tensor of the stellar and dark matter distributions, defining a {\it shape parameter} $s = 1 - \lambda_{2}/\lambda_{1}$ ($\lambda_{1} \geq \lambda_{2}$; see Section~\ref{sec:shapemethod}). We averaged $s$ over the range 1-5$R_e$, where $R_e$ is the effective radius of the light profile.

>>>>

\begin{abstract}
We measure the projected density profile, shape and alignment of the stellar and dark matter mass distribution in 11 strong-lens galaxies. We find that the projected dark matter density profile -- under the assumption of a Chabrier stellar initial mass function -- shows significant variation from galaxy to galaxy. Those with an outermost image beyond $\sim 10$\,kpc are very well fit by a projected NFW profile; those with images within 10\,kpc appear to be more concentrated than NFW, as expected if their dark halos contract due to baryonic cooling. We find that over several half-light radii, the dark matter haloes of these lenses are rounder than their stellar mass distributions. While the haloes are never more elliptical than $s_{dm} = 0.3$ (where $s$ is defined such that $s=0$ is spherical and $s=1$ is infinitely flattened), their stars can extend to $s_* > 0.3$. Galaxies with high dark matter ellipticity and weak external shear show strong alignment between light and dark; those with strong shear ($\gamma \gtrsim 0.1$) can be highly misaligned. This is reassuring since isolated misaligned galaxies are expected to be unstable. Our results provide a new constraint on galaxy formation models. For a given cosmology, these must explain the origin of both very round dark matter haloes and misaligned strong-lens systems.
{\bf JR. I hope you agree with my analysis of the radial profiles. I've spotted a very cute little result here already and thought we should highlight it. I discuss it further in the results and conclusions. Also, we now define $s$ to be the `ellipticity'. But later on when discussing weak lensing we define a different ellipticity parameter $e$. The referee took us to task on this also I think. Are there really two different definitions of ellipticity in the literature? If so, perhaps we should clarify this up-front when we define $s$. Alternatively, if $s$ is unique to our paper then why not just change $s \rightarrow e$ throughout? Or we could call our $s$ parameter the `shape parameter' and $e$ the `ellipticity'. I have no strong view on which is best, but it's a bit confusing as it is right now}.
\end{abstract}

@@

Our lens sample is small -- just 11 strong lensing systems -- yet it is interesting that we find no lens with strong misalignment and weak external shear. This is perhaps to be expected. \citet{1979ApJ...233..872H} were the first to study the stability of orbits within triaxial figures, likely the physical situation in real elliptical galaxies. They found that while stable periodic orbits exist in all three symmetry planes of a triaxial figure, tube orbits about the intermediate axis are unstable. \citet{1985MNRAS.215..731D} studied the case of orbits misaligned to all three symmetry planes. They found that while misaligned stable orbits can be found, they exit only for limited ranges of the Hamiltonian. \citet{1988A&A...206..269M} extended these results to rotating triaxial figures showing that once rotation is included such orbits become even rarer. These early results are supported by more recent numerical work. \citet{2013MNRAS.434.2971D} show that discs can never form perpendicular to the intermediate axis, in good agreement with earlier work by \citet{1979ApJ...233..872H}. However, they find that discs can survive off-axis if gas is present, since gas cooling can maintain the misalignment. \citet{2015MNRAS.452.4094D} go on to show that if this gas supply is switched off (as is the case for giant elliptical galaxies), then continuing collisionless minor mergers/interactions bring the disc rapidly into alignment with its host halo. Taken together, these results suggest that alignment between light and dark should be the norm for isolated gas-free triaxial galaxies.

\section{Conclusions}\label{sec:conclusions}
We have measured the projected radial density profile, shape and alignment of the stellar and dark matter distributions in 11 lens systems using a new non-parametric lens tool, \Glass. We focussed on lenses that have either time delay data or stellar mass maps that contribute significantly to the central potential. In a previous paper, we showed that data of this quality are required to determine the projected shape of dark matter haloes \citep{2014MNRAS.445.2181C}. We measured the shape and alignment using the eigenvalues $\lambda_i$ and eigenvectors of the 2D moment of inertia tensor of the stellar and dark matter distributions, defining a {\it shape parameter} $s = 1 - \lambda_{2}/\lambda_{1}$ ($\lambda_{1} \geq \lambda_{2}$; see Section~\ref{sec:shapemethod}). We averaged $s$ over the range 1-5$R_e$, where $R_e$ is the effective radius of the light profile.
